---
title: "Reproducible Research: Peer Assessment 1"
author: "niknik324"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

Getting data from activity.csv
```{r, echo=TRUE}
activity <- read.csv('activity.csv')
```

No need to transform.

## What is mean total number of steps taken per day?

First - split and sum data.  Days without any activity are excluded.
```{r, echo=TRUE}
perday <- aggregate(steps ~ date, data = activity, sum)
```


#####Now let's make our histogram.
```{r, echo=TRUE}
hist(perday$steps, xlab = 'steps per day', main = 'Histogram of steps per day')
```


#####**Mean** and **median**
```{r, echo=TRUE}
perdayMean <- mean(perday$steps)
perdayMedian <- median(perday$steps)
```

These are: 

**Mean:** `r as.integer(perdayMean)`

**Median:** `r perdayMedian`


## What is the average daily activity pattern?
First calculate mean steps per interval
```{r, echo=TRUE}
meanPerInterval <- aggregate(steps ~ interval, data = activity, mean)
```

Now - plot
First calculate mean steps per interval
```{r, echo=TRUE}
plot(meanPerInterval, type = 'l')
```

Most active interval is "`r meanPerInterval[meanPerInterval$steps == max(meanPerInterval$steps),1]`"
```{r, echo=TRUE}
meanPerInterval[meanPerInterval$steps == max(meanPerInterval$steps),1]
```


## Imputing missing values
There are `sum(is.na(activity$steps))` intevals with no infomation about steps available
```{r, echo=TRUE}
sum(is.na(activity$steps))
```


I will replace ol NA with mean value for that interval. And yes, I'm doing it not the easiest way.
```{r, echo=TRUE}
activityFIXED <- activity
for (row in 1:17568) { 
    if (is.na(activityFIXED[row,]$steps)) {
        activityFIXED[row,]$steps <- meanPerInterval[2][meanPerInterval[1]==activityFIXED[row,]$interval]
    }
}
head(activityFIXED)
```

Here is my new Histogram

```{r, echo=TRUE}
perdayF <- aggregate(steps ~ date, data = activityFIXED, sum)
hist(perdayF$steps, xlab = 'steps per day', main = 'Histogram of steps per day')

perdayMeanF <- mean(perdayF$steps)
perdayMedianF <- median(perdayF$steps)
```


**Mean:** `r as.integer(perdayMeanF)` 

**Median:** `r as.integer(perdayMedianF)`


## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
activityFIXED$weekday <- weekdays(as.Date(activityFIXED$date))
for (row in 1:17568) {if (activityFIXED[row,]$weekday == "Sunday" | activityFIXED[row,]$weekday == "Saturday"|activityFIXED[row,]$weekday == "воскресенье"|activityFIXED[row,]$weekday == "суббота") { activityFIXED[row,]$weekday <- 'weekend'} else {activityFIXED[row,]$weekday <- 'weekday'} }

meanPerIntervalWD <- aggregate(steps ~ interval, data = activityFIXED[activityFIXED$weekday == 'weekday',], mean)
meanPerIntervalWE <- aggregate(steps ~ interval, data = activityFIXED[activityFIXED$weekday == 'weekend',], mean)

par(mfcol=c(2,1))
plot(meanPerIntervalWD, type = 'l')
plot(meanPerIntervalWE, type = 'l')

```

Yes, they are different.